---
title: "Euler Problem 3: Largest Prime Factor"
author: "Benjamin Chan"
date: 2017-10-08 19:04:09
categories: ["Project_Euler"]
tags: ["R", "Python", "Project_Euler"]
description: "How to find the largest prime factor of a number"
---

<!DOCTYPE html PUBLIC "-//W3C//DTD HTML 3.2 Final//EN">
<html>
<head>
<title>Mathedemo</title>
<script type="text/x-mathjax-config">
  MathJax.Hub.Config({tex2jax: {inlineMath: [['$','$'], ['\\(','\\)']]}});
</script>
<script type="text/javascript" async
  src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.2/MathJax.js?config=TeX-MML-AM_CHTML">
</script>
</head>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the previous post, I introduced the the Sieve of Eratosthenes to find out all prime numbers with in a certain number. In this post, I'm going to use it on the Euler Problem 3.

## Euler Problem 3

The prime factors of $13195$ are $5, 7, 13$ and $29$. 

What is the largest prime factor of the number $600851475143$?

---

I use `all_prime()` function to find out all prime numbers less than or equal to the largest integer less than square root of $n$. Because we are looking for the largest **prime factor** of $n$ but not the largest prime number within $n$, we don't need to generate all primes within $n$. After we have the list of primes, we can loop through it to find all prime factors. Please note after the `for` loop, if `n / j` $> 1$, `n / j` is also a prime factor of $n$.

### Solutions

#### Solution in R
```{r}
all_primes <- function(n) {
  prime <- rep(TRUE, n)  # Intitialize a vector and all set to TRUE
  prime[1] <- FALSE  # Remove 1 from output
  
  for (i in 2:sqrt(n)) {
    if (prime[i]) prime[seq(i ^ 2, n, i)] <- FALSE  # Mark FALSE if composite index.
  }
  
  which(prime)  # Return indices of TRUE
}

prime_factors <- function(n) {
  factors <- c()  # Create a empty vector to store prime factors
  prime_numbers <- all_primes(floor(sqrt(n)))  # Find out all primes no greater than square root of n
  d <- which(n %% prime_numbers == 0)  # Identify the n's prime factors
  
  if (length(d) == 0) return(n)  # Return n if n has no prime divisors
  
  for (j in prime_numbers[d]) {  # Loop through all prime numbers
    while (n %% j == 0) {
      factors <- c(factors , j)  # Append the list of prime factors
      n <- n / j
    }
  }
  if (n > 1) factors <- c(factors, n)  # This new n is also a prime factor of the original n
  
  return(factors)
}

max(prime_factors(600851475143))
```

#### Solution in Python
```{python, eval = FALSE}
import math
import numpy as np

def all_primes(n):
    prime = np.array([True] * (n + 1))  # Intitialize a vector and all set to TRUE
    prime[0] = prime[1] = False  # Remove 0 and 1 from output

    for i in range(2, int(math.ceil(math.sqrt(n))) + 1):
        if prime[i]:
            prime[range(i ** 2, n + 1, i)] = False  # Mark FALSE if composite index.

    return(np.where(prime == 1))  # Return indices of TRUE

def prime_factors(n):
    factors = []  # Create a empty list to store prime factors
    prime_numbers = all_primes(math.floor(math.sqrt(n)))  # Find out all primes no greater than square root of n
    d = np.where(n % np.array(prime_numbers) == 0)[1]  # Create a numpy.ndarray to identify the n's prime factors

    if len(d) == 0:
        return(n)  # Return n if n has no prime divisors

    for j in prime_numbers[0][d]:  # Loop through all prime numbers in a numpy.ndarray
        while n % j == 0:
            factors.append(j)  # Append the list of prime factors
            n /= j

    if n > 1:
        factors.append(n)  # This new n is also a prime factor of the original n

    return(factors)

print(max(prime_factors(600851475143)))
```

To test my solutions, we can use `primeFactors` function in **numbers** package.
```{r}
library(numbers)
max(primeFactors(600851475143))
```