---
title: "Euler Problem 4: Largest palindrome product"
author: "Benjamin Chan"
date: 2017-10-15 16:50:55
categories: ["Project_Euler"]
tags: ["R", "Python", "Project_Euler"]
description: "Euler Problem 4: Largest palindrome product"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Euler Problem 4

A palindromic number reads the same both ways. The largest palindrome made from the product of two 2-digit numbers is 9009 = 91 × 99. Find the largest palindrome made from the product of two 3-digit numbers.

The solution is to search for palindrome number starting from the largest number. We check if the product of two numbers is a palindrome and replace k with it if it's larger than k. We return k, the largest palindrome, after the loops.

### Solutions

#### Solution in R
```{r}
is_palindrome <- function(x) {
  s <- as.character(x)
  s_reverse <- paste(rev(unlist(strsplit(s, ""))), collapse = "")
  return(s == s_reverse)
}

n = 0
m = 0
k = 0

for (i in 999:100) {
  for (j in 999:100) {
    number <- i * j
    if (is_palindrome(number) & number > k) {
      n = i
      m = j
      k = number
    }
  }
}

paste(n, "x", m, "=", k)
```

#### Solution in Python
```{python}
def is_palindrome(x):
    s = str(x)
    return s == s[::-1]

n = 0
m = 0
k = 0

for i in range(999, 99, -1):
    for j in range(999, 99, -1):
        number = i * j
        if is_palindrome(number) and number > k:
            n = i
            m = j
            k = number
        
print("%d x %d = %d" %(n, m, k))
```