---
title: 'Euler Problem 7: 10001st prime'
author: "Benjamin Chan"
date: 2017-10-21 15:55:11
categories: ["Project_Euler"]
tags: ["R", "Python", "Project_Euler"]
description: "How to find the Nth prime"
---

<!DOCTYPE html PUBLIC "-//W3C//DTD HTML 3.2 Final//EN">
<html>
<head>
<title>Mathedemo</title>
<script type="text/x-mathjax-config">
  MathJax.Hub.Config({tex2jax: {inlineMath: [['$','$'], ['\\(','\\)']]}});
</script>
<script type="text/javascript" async
  src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.2/MathJax.js?config=TeX-MML-AM_CHTML">
</script>
</head>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You might ask why I skip Euler Problem 6. It's because Euler Problem 6 is so easy that I don't want to write a blog about its solution. In this post, we will discuss Euler Problem 7.

## Euler Problem 7

By listing the first six prime numbers: 2, 3, 5, 7, 11, and 13, we can see that the 6th prime is 13. 

What is the 10001st prime number?

---

### Solution

To generalize this problem, we can ask "What is the $N^{th}$ prime number?". In Euler Problem 3, we built `all_primes()` function to find out how many prime numbers less than or equal to a given number. We, however, cannot use this function to find out the target prime because we don't know which number is bigger than the target.

My strategy is following:

1.Create a list `primes` to save each prime.

2.`primes` is initialized with first prime.

3.Check if the length of `primes` is less than number $N$. If so, do step 4 and 5.

4.Start with 3 and loop through odd numbers. Check if it's a prime by the helper function `is_prime()`.

5.Add the number to `primes` if `is_prime()` returns true.

6.Loop stops if the length of `primes` equal to $N$.


#### Solution in R
```{r}
is_prime <- function(n) {
  if (n == 3) {
    return(TRUE)
  }
  
  if (n %% 2 == 0) {  # Even numbers bigger than 2 are no primes
    return(FALSE)
  }
  
  for (i in seq(3, n ^ 0.5 + 1, 2)) {  # Only loop through odd numbers
    if (n %% i == 0) {
      return(FALSE)
    }
  }
  
  return(TRUE)
}


get_prime2 <- function(x) {
  primes <- c(2)
  m <- 3
  
  while (length(primes) < x) {  # Loop until we get x amount of primes
    if (is_prime(m)) {
      primes <- c(primes, m)
    }
    
    m <- m + 2  # Only loop through odd numbers
  }
  
  primes[length(primes)]
}


get_prime2(10001)
```

#### Solution in Python
```{python}
def is_prime(n):
    if n == 3:
        return True

    if n % 2 == 0:  # Even numbers bigger than 2 are no primes
        return False

    for i in range(3, int(n ** 0.5) + 1, 2):  # Only loop through odd numbers
        if n % i == 0:
            return False

    return True


def get_prime(x):
    primes = [2]
    m = 3

    while len(primes) < x:  # Loop until we get x amount of primes
        if is_prime(m):
            primes.append(m)
        m += 2  # Only loop through odd numbers

    print(primes[-1])


get_prime(10001)
```

